---
title: "PostHog Engineering Impact — Methodology & Data"
output:
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    self_contained: true
params:
  data_dir: "raw-data"
  date_start: "2025-12-02"
  date_end: "2026-03-02"
  dashboard_url: "https://avery2.github.io/weave-takehome/"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(jsonlite)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(plotly)
library(DT)
library(crosstalk)
```

[← Back to Dashboard](`r params$dashboard_url`)

## 1. Data Overview

```{r load-data}
commits <- fromJSON(file.path(params$data_dir, "commits.json"), flatten = TRUE)
issues <- fromJSON(file.path(params$data_dir, "issues.json"), flatten = TRUE)
comments <- fromJSON(file.path(params$data_dir, "issue_comments.json"), flatten = TRUE)

has_prs <- file.exists(file.path(params$data_dir, "prs.json"))
if (has_prs) {
  prs <- fromJSON(file.path(params$data_dir, "prs.json"), flatten = TRUE)
} else {
  prs <- data.frame()
}
```

```{r overview-table}
overview <- tibble(
  Metric = c("Repository", "Date range", "Total commits", "Unique commit authors",
             "Merged PRs", "Closed issues", "Issue comments"),
  Value = c(
    "PostHog/posthog",
    paste(params$date_start, "to", params$date_end),
    format(nrow(commits), big.mark = ","),
    format(n_distinct(commits$author_email), big.mark = ","),
    if (has_prs) format(nrow(prs), big.mark = ",") else "Not yet collected",
    format(nrow(issues), big.mark = ","),
    format(nrow(comments), big.mark = ",")
  )
)
knitr::kable(overview, align = "lr")
```

## 2. Signal Distributions

These raw signal distributions show the spread of activity across contributors before any scoring is applied.

### Commits per Author

```{r commits-per-author}
commits_by_author <- commits %>%
  filter(!is_merge) %>%
  group_by(author_name, author_email) %>%
  summarise(
    commit_count = n(),
    total_additions = sum(total_additions),
    total_deletions = sum(total_deletions),
    files_touched = sum(lengths(files)),
    .groups = "drop"
  ) %>%
  arrange(desc(commit_count))

sd_commits <- SharedData$new(commits_by_author, key = ~author_email)

bscols(
  widths = c(6, 6),
  plot_ly(sd_commits, x = ~commit_count, type = "histogram",
          marker = list(color = "#4f46e5")) %>%
    layout(title = "Distribution of Commits per Author",
           xaxis = list(title = "Commits"), yaxis = list(title = "Authors")),
  datatable(sd_commits,
            colnames = c("Author", "Email", "Commits", "Additions", "Deletions", "Files"),
            options = list(pageLength = 10, dom = "ftp"),
            rownames = FALSE) %>%
    formatRound(c("total_additions", "total_deletions", "files_touched"), digits = 0)
)
```

### Commit Activity Over Time

```{r commits-over-time}
commits_daily <- commits %>%
  mutate(date = as.Date(timestamp)) %>%
  group_by(date) %>%
  summarise(count = n(), .groups = "drop")

plot_ly(commits_daily, x = ~date, y = ~count, type = "bar",
        marker = list(color = "#4f46e5")) %>%
  layout(title = "Daily Commit Volume",
         xaxis = list(title = ""), yaxis = list(title = "Commits"))
```

```{r pr-section, eval=has_prs, results='asis'}
cat("\n### PRs per Author\n")
```

```{r pr-dist, eval=has_prs}
prs_by_author <- prs %>%
  group_by(author) %>%
  summarise(
    pr_count = n(),
    total_additions = sum(additions),
    total_deletions = sum(deletions),
    median_cycle_days = median(as.numeric(
      difftime(ymd_hms(merged_at), ymd_hms(created_at), units = "days")
    ), na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(pr_count))

sd_prs <- SharedData$new(prs_by_author, key = ~author)

bscols(
  widths = c(6, 6),
  plot_ly(sd_prs, x = ~pr_count, type = "histogram",
          marker = list(color = "#059669")) %>%
    layout(title = "PRs Merged per Author",
           xaxis = list(title = "PRs Merged"), yaxis = list(title = "Authors")),
  datatable(sd_prs,
            colnames = c("Author", "PRs", "Additions", "Deletions", "Median Cycle (days)"),
            options = list(pageLength = 10, dom = "ftp"),
            rownames = FALSE) %>%
    formatRound(c("total_additions", "total_deletions", "median_cycle_days"), digits = 1)
)
```

```{r pr-cycle-time, eval=has_prs}
prs_with_cycle <- prs %>%
  mutate(cycle_hours = as.numeric(
    difftime(ymd_hms(merged_at), ymd_hms(created_at), units = "hours")
  )) %>%
  filter(cycle_hours >= 0, cycle_hours < 720) # exclude outliers > 30 days

plot_ly(prs_with_cycle, x = ~cycle_hours, type = "histogram",
        marker = list(color = "#059669")) %>%
  layout(title = "PR Cycle Time Distribution (hours, < 30 days)",
         xaxis = list(title = "Hours to Merge"), yaxis = list(title = "PRs"))
```

### Issues Closed

```{r issues-dist}
issues_by_creator <- issues %>%
  group_by(creator) %>%
  summarise(issues_closed = n(), .groups = "drop") %>%
  arrange(desc(issues_closed))

plot_ly(issues_by_creator %>% head(30), x = ~reorder(creator, -issues_closed),
        y = ~issues_closed, type = "bar",
        marker = list(color = "#d97706")) %>%
  layout(title = "Issues Closed by Creator (top 30)",
         xaxis = list(title = "", tickangle = -45),
         yaxis = list(title = "Issues Closed"))
```

## 3. Dimension Breakdowns

Scores are computed across four dimensions. Each signal is normalized to a 0–100 scale using percentile rank within the contributor pool before combining.

```{r compute-dimensions, eval=has_prs}
# Build per-engineer signal table
# Start with commits (non-merge)
eng_commits <- commits %>%
  filter(!is_merge) %>%
  group_by(author = author_name) %>%
  summarise(
    commit_count = n(),
    lines_added = sum(total_additions),
    lines_deleted = sum(total_deletions),
    .groups = "drop"
  )

# PR signals
eng_prs <- prs %>%
  mutate(cycle_days = as.numeric(
    difftime(ymd_hms(merged_at), ymd_hms(created_at), units = "days")
  )) %>%
  group_by(author) %>%
  summarise(
    pr_count = n(),
    median_cycle_days = median(cycle_days, na.rm = TRUE),
    total_pr_additions = sum(additions),
    total_pr_deletions = sum(deletions),
    .groups = "drop"
  )

# Review signals (from PR reviews data)
reviews_flat <- prs %>%
  select(number, reviews) %>%
  unnest(reviews, names_sep = "_") %>%
  filter(!is.na(reviews_author))

eng_reviews <- reviews_flat %>%
  group_by(author = reviews_author) %>%
  summarise(
    reviews_given = n(),
    approvals_given = sum(reviews_state == "APPROVED"),
    changes_requested = sum(reviews_state == "CHANGES_REQUESTED"),
    review_comments = sum(reviews_state == "COMMENTED"),
    .groups = "drop"
  )

# File classification: user-facing = frontend/*, *.tsx, *.css, templates/
classify_user_facing <- function(files_list) {
  if (is.null(files_list) || length(files_list) == 0) return(0)
  paths <- if (is.data.frame(files_list)) files_list$path else files_list
  uf <- grepl("^frontend/|\\.(tsx|jsx|css|scss)$|templates/", paths, ignore.case = TRUE)
  mean(uf, na.rm = TRUE)
}

pr_files <- prs %>%
  rowwise() %>%
  mutate(user_facing_pct = classify_user_facing(files)) %>%
  ungroup()

eng_uf <- pr_files %>%
  group_by(author) %>%
  summarise(user_facing_pct = mean(user_facing_pct, na.rm = TRUE), .groups = "drop")

# Label-based signals
eng_labels <- prs %>%
  rowwise() %>%
  mutate(
    is_bug = any(grepl("bug|fix|hotfix", labels, ignore.case = TRUE)),
    is_feature = any(grepl("feat|feature|enhancement", labels, ignore.case = TRUE))
  ) %>%
  ungroup() %>%
  group_by(author) %>%
  summarise(
    bug_prs = sum(is_bug),
    feature_prs = sum(is_feature),
    .groups = "drop"
  )

# Issues closed (map by username since issue creator ≠ always the closer)
eng_issues <- issues %>%
  group_by(author = creator) %>%
  summarise(issues_closed = n(), .groups = "drop")

# Join everything
engineers <- eng_prs %>%
  full_join(eng_commits, by = "author") %>%
  full_join(eng_reviews, by = "author") %>%
  full_join(eng_uf, by = "author") %>%
  full_join(eng_labels, by = "author") %>%
  full_join(eng_issues, by = "author") %>%
  mutate(across(where(is.numeric), ~replace_na(., 0)))

# Percentile rank normalization (0-100)
prank <- function(x) {
  r <- rank(x, ties.method = "average", na.last = "keep")
  100 * (r - 1) / (max(r, na.rm = TRUE) - 1)
}

# Dimension scores
engineers <- engineers %>%
  mutate(
    # Delivery: PR count, commit count, cycle time (inverse), issues closed
    d_delivery = (prank(pr_count) + prank(commit_count) +
                  prank(-median_cycle_days) + prank(issues_closed)) / 4,

    # Scope: user-facing %, feature PR ratio, total additions
    d_scope = (prank(user_facing_pct) + prank(feature_prs) +
               prank(total_pr_additions)) / 3,

    # Collaboration: reviews given, review diversity (approvals + changes + comments)
    d_collaboration = (prank(reviews_given) + prank(changes_requested) +
                       prank(review_comments)) / 3,

    # Stability: low churn = low deletions relative to additions
    churn_rate = ifelse(lines_added > 0, lines_deleted / lines_added, 0),
    d_stability = prank(-churn_rate)
  )

# Composite (equal weights by default)
engineers <- engineers %>%
  mutate(
    composite = (d_delivery + d_scope + d_collaboration + d_stability) / 4
  ) %>%
  arrange(desc(composite))
```

```{r dimension-table, eval=has_prs}
engineers %>%
  select(author, d_delivery, d_scope, d_collaboration, d_stability, composite) %>%
  head(30) %>%
  datatable(
    colnames = c("Engineer", "Delivery", "Scope", "Collaboration", "Stability", "Composite"),
    options = list(pageLength = 15, dom = "ftp"),
    rownames = FALSE
  ) %>%
  formatRound(c("d_delivery", "d_scope", "d_collaboration", "d_stability", "composite"), digits = 1)
```

```{r dimension-radar-top5, eval=has_prs, fig.width=10, fig.height=6}
top5 <- engineers %>% head(5)

radar_data <- top5 %>%
  select(author, d_delivery, d_scope, d_collaboration, d_stability) %>%
  pivot_longer(-author, names_to = "dimension", values_to = "score") %>%
  mutate(dimension = gsub("d_", "", dimension) %>% tools::toTitleCase())

plot_ly(type = "scatterpolar", mode = "lines+markers", fill = "toself") %>%
  add_trace(
    r = ~score, theta = ~dimension,
    data = radar_data %>% filter(author == top5$author[1]),
    name = top5$author[1]
  ) %>%
  add_trace(
    r = ~score, theta = ~dimension,
    data = radar_data %>% filter(author == top5$author[2]),
    name = top5$author[2]
  ) %>%
  add_trace(
    r = ~score, theta = ~dimension,
    data = radar_data %>% filter(author == top5$author[3]),
    name = top5$author[3]
  ) %>%
  add_trace(
    r = ~score, theta = ~dimension,
    data = radar_data %>% filter(author == top5$author[4]),
    name = top5$author[4]
  ) %>%
  add_trace(
    r = ~score, theta = ~dimension,
    data = radar_data %>% filter(author == top5$author[5]),
    name = top5$author[5]
  ) %>%
  layout(title = "Top 5 Engineers — Dimension Profiles",
         polar = list(radialaxis = list(range = c(0, 100))))
```

## 4. Composite Scores & Archetypes

```{r archetypes, eval=has_prs}
assign_archetype <- function(row) {
  scores <- c(
    firefighter = as.numeric(row["bug_prs"]) / max(engineers$bug_prs, 1) * 100,
    scaffolder = (1 - as.numeric(row["user_facing_pct"])) * as.numeric(row["d_delivery"]),
    ux_driver = as.numeric(row["user_facing_pct"]) * 100,
    reviewer = as.numeric(row["d_collaboration"]),
    integrator = (as.numeric(row["d_delivery"]) + as.numeric(row["issues_closed"]) /
                  max(engineers$issues_closed, 1) * 100) / 2
  )
  names(which.max(scores))
}

archetype_labels <- c(
  firefighter = "\U0001f525 The Firefighter",
  scaffolder = "\U0001f3d7\ufe0f The Scaffolder",
  ux_driver = "\U0001f3a8 The UX Driver",
  reviewer = "\U0001f52d The Reviewer",
  integrator = "\U0001f9e9 The Integrator"
)

archetype_descriptions <- c(
  firefighter = "Closes the most bugs and hotfixes",
  scaffolder = "Builds foundational infrastructure",
  ux_driver = "Owns the user-facing surface",
  reviewer = "Unblocks the team through reviews",
  integrator = "Cross-cutting work across many areas"
)

top5 <- engineers %>% head(5)
top5$archetype_key <- apply(top5, 1, assign_archetype)
top5$archetype <- archetype_labels[top5$archetype_key]
top5$archetype_desc <- archetype_descriptions[top5$archetype_key]
```

```{r top5-table, eval=has_prs}
top5 %>%
  select(archetype, author, d_delivery, d_scope, d_collaboration, d_stability, composite) %>%
  datatable(
    colnames = c("Archetype", "Engineer", "Delivery", "Scope", "Collaboration", "Stability", "Composite"),
    options = list(pageLength = 5, dom = "t"),
    rownames = FALSE
  ) %>%
  formatRound(c("d_delivery", "d_scope", "d_collaboration", "d_stability", "composite"), digits = 1)
```

```{r top5-detail, eval=has_prs, results='asis'}
for (i in 1:nrow(top5)) {
  eng <- top5[i, ]
  cat(sprintf("\n### %s — %s\n\n", eng$archetype, eng$author))
  cat(sprintf("*%s*\n\n", eng$archetype_desc))
  cat(sprintf("- **PRs merged:** %d | **Commits:** %d | **Reviews given:** %d\n",
              eng$pr_count, eng$commit_count, eng$reviews_given))
  cat(sprintf("- **Median cycle time:** %.1f days\n", eng$median_cycle_days))
  cat(sprintf("- **User-facing file %%:** %.0f%%\n\n", eng$user_facing_pct * 100))
}
```

## 5. Methodology Notes

### How dimensions are computed

Each dimension aggregates multiple raw signals. Before combining, each signal is normalized to a **0–100 percentile rank** within the contributor pool. This prevents any single signal (e.g., lines of code) from dominating.

| Dimension | Signals used | Notes |
|-----------|-------------|-------|
| **Delivery** | PR count, commit count, inverse cycle time, issues closed | Cycle time is inverted so faster = higher score |
| **Scope** | User-facing file %, feature PR count, total additions | "User-facing" = files in `frontend/`, `.tsx/.jsx/.css` files |
| **Collaboration** | Reviews given, changes requested, review comments | Measures review depth, not just count |
| **Stability** | Inverse churn rate (deletions/additions) | Lower churn = more stable contributions |

### Composite score

The composite is a simple average of the four dimension scores (equal 25% weights). In the dashboard UI, weights are adjustable via sliders.

### Archetype assignment

Each of the top 5 engineers is assigned one primary archetype based on which contribution pattern they score highest in. The archetype is **descriptive** — it says what kind of work someone does, not how good they are.

### Known limitations

- **Churn rate** is a rough proxy for stability — it doesn't distinguish intentional refactors from bugs
- **User-facing classification** uses file paths only — misses backend changes that directly affect UX
- **Review quality** is approximated by comment count, not actual review thoroughness
- **Bot activity** is included in raw data and filtered only if detected by GitHub username patterns
- **Issue attribution** uses the issue creator, not necessarily who resolved it

## 6. Open Questions

What this analysis **doesn't yet capture**:

- **Review quality**: Are reviews substantive or rubber-stamp approvals?
- **Product impact**: Did the shipped code actually improve user metrics?
- **Cross-team work**: Collaboration beyond the posthog repo
- **Non-GitHub signals**: Slack discussions, design docs, mentoring, incident response
- **Revert rate**: How often are PRs reverted after merge?
- **Deploy frequency**: How often do changes reach production?

---

*Data collected from [PostHog/posthog](https://github.com/PostHog/posthog) for `r params$date_start` to `r params$date_end`.*

[← Back to Dashboard](`r params$dashboard_url`)
